The Unintended Consequences of Energy Reduction

### 2024-02-14 paper/paper.qmd
Wrote introduction with citations and references

# Introduction

![Figure 1](../other/figures/figure1.png)

The 2011 Fukushima disaster changed the relationship Japanese citizens had with nuclear energy. From the 160, 000 Fukushima residents that had to evacuate the city to the citizens who may not have been affected directly, the public as a whole turned their backs against the entire industry [science direct].

Before the accident, nuclear energy made up 25% of Japan’s energy market. After the accident, the market share stayed under 1.7% until 2016. The Democratic Party of Japan itself took initiative and implemented a new energy policy to “phase out nuclear power,” [intchopen] despite formerly having close ties to the industry in attempts to reduce CO2 pollution. [enpolicy one year later]

Fossil fuels grew to dominate Japan’s energy market as many reactors were shut down. In 2011 itself, oil usage rose by around 85% while natural gas increased by 25%. [enpolicy one year later]

This change had a huge impact on the price and consumption of electricity. Relative to the mean price of electricity in 2011, by 2013 the price of electricity had increased to over 10% and almost reached an increase of 20% by 2014. Concurrently, electricity consumption itself kept decreasing year after year as it became less affordable. [figure 1]

A good reason for this market shift is that most of the fossil fuels that Japan utilizes are exported from other countries, which makes them expensive. [enpolicy one year later] The lack of nuclear power also reduced more affordable options for consumers.

The trend of increasing electricity prices and decreasing electricity consumption may persist into the future. The current government of Japan reversed the nuclear energy policy, but low public demand for nuclear energy still persists. Thus the prices will stay high as people continue to consume fossil fuels. There are also barriers to nuclear power that help to persist this trend. One barrier is the Nuclear Regulation Authority which created tougher safety regulations for reactors. Another are injunctions that citizens can bring forth to prevent the operation of nuclear reactors. After the Fukushima tragedy many courts in Japan may grant these injunctions for safety reasons.  [intchopen]

![Figure 5](../other/figures/figure5.png)
After the Fukushima incident, the government tried to ensure sufficient backup capacity to avoid power blackouts. While other factors could affect the saving target, the primary determinant of the saving target was a region’s dependence on nuclear power before the accident.(Kim, et al.) This caused more health issues for individuals as they were encouraged to not use AC and other cooling devices during hotter days. Previous to the implementation of the energy policy, throughout the period between 1960 and 2004, the mortality rate at over 32.2 degrees Celsius has dropped tremendously while that of between 26.7-32.3 degrees Celsius remained consistent. However, it becomes evident that after the policy implementation, at similar ranges of 25-30 degrees Celsius and over 30 degrees Celsius produced a higher mortality rate. [figure 5]

>>>>

# Introduction
The 2011 Fukushima nuclear disaster changed the relationship Japanese citizens have with nuclear energy. From the 160,000 Fukushima residents that had to evacuate the city to the citizens who may not have been affected directly, the public as a whole turned their backs against the entire nuclear industry [@scienceDirect]. However, we find that Japan’s halt in nuclear energy generation is not without its fatal consequences.

The nuclear disaster was triggered by an earthquake, generating tsunami waves that flooded the Fukushima Daiichi plant and shut off power to its cooling systems, causing the meltdown of three reactor cores and radiation release [@britannica]. Fortunately, there was no death toll as a result of radiation exposure; however the aftermath of the accident did not leave the Japanese public unscathed. Thousands of residents’ mental and physical health deteriorated as they were forced to evacuate and be permanently relocated [@CDC]. According to the CDC, the disaster triggered an increase in deaths among elderly citizens placed in temporary housing and a rise in post-traumatic stress disorder rates among those affected [@CDC]. The public lost trust in nuclear energy and demanded the use of nuclear power plants be eradicated from Japan [@intechopen].

As a response to the growing public outcry, the government of Japan took swift action to significantly reduce nuclear power generation by implementing a new energy policy to eliminate nuclear power by 2030 [@intechopen]. Before the accident, nuclear energy comprised 25% of Japan’s energy market [@intechopen]. After the accident, the market share stayed under 1.7% until 2016 [@intechopen]. The significant and abrupt reduction in overall power generation caused Japan to steer towards potential countrywide power shortages [@guardian]. To combat this, the government pushed a power saving campaign to the public [@guardian]. In fear of blackouts, energy saving was quickly adopted by Japanese citizens, resulting in a considerable decrease in electricity consumption [@intechopen].

The energy saving campaign was not the sole factor in the sharp decline in energy consumption as the price for electricity also surged. As many nuclear reactors were shut down, fossil fuels grew to dominate Japan’s energy market. In 2011, oil usage rose by an estimated 85% while natural gas increased by 25% [@energyPolicyYearLater]. This shift to predominate fossil fuel usage had a substantial impact on the price and consumption of electricity as the majority of fossil fuels utilized in Japan are imported from other countries, resulting in expensive prices [@energyPolicyYearLater]. Relative to the mean price of electricity in 2011, the price of electricity had increased by over 10% in 2013 and further increased by nearly 20% in 2014 [@energyPolicyYearLater]. Consequently, electricity consumption decreased each year as the commodity became less affordable. The Japanese government was successful in reducing energy consumption; however, their energy-saving policies and campaigns cause potential detrimental consequences to the public.

In this paper, we follow He and Tanaka’s paper to produce a reproduction of their findings and discuss further about Japan’s future as their government reverses their stance on nuclear energy and begins restarting the use of old power plants as a response to fossil fuel shortages and the public pressure to reduce the country’s carbon footprint. We analyze the major reduction in energy consumption from 2005 to 2015 caused by the decline in nuclear energy usage, the sharp increase in electricity prices and the energy saving campaign pushed by the Japanese government to investigate their impact on mortality rate in high temperatures. We find that the mortality rate increases as energy consumption decreases. An important finding is that air conditioner usage plays a vital role in preventing deaths related to high temperatures. Therefore, the pressure to conserve energy and hence not utilize air conditioners, increases the mortality rate in Japan.

The remainder of this paper is structured as follows: the Data section details the collection
and processing of nuclear energy production, energy consumption and energy price, and temperature-mortality data. Additionally, the section includes discussion on the measurement of the original dataset. The following Results section displays and analyzes tables
and figures that illustrate the trends found in the data. The Discussion section draws conclusions from those tables and figures and details the importance of our findings and weaknesses of our paper.

![Figure 1](../other/figures/figure1.png)
![Figure 5](../other/figures/figure5.png)
After the Fukushima incident, the government tried to ensure sufficient backup capacity to avoid power blackouts. While other factors could affect the saving target, the primary determinant of the saving target was a region’s dependence on nuclear power before the accident.(Kim, et al.) This caused more health issues for individuals as they were encouraged to not use AC and other cooling devices during hotter days. Previous to the implementation of the energy policy, throughout the period between 1960 and 2004, the mortality rate at over 32.2 degrees Celsius has dropped tremendously while that of between 26.7-32.3 degrees Celsius remained consistent. However, it becomes evident that after the policy implementation, at similar ranges of 25-30 degrees Celsius and over 30 degrees Celsius produced a higher mortality rate. [figure 5]
